Predicting the effects of character displacement on food-web dynamics

### 2019-11-18 manuscript_ECD_model.Rmd
added Zenodo DOI to manuscript.

*Data accessibility*: No new data were used. All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model)<!-- and have been archived with Zenodo (INSERT DOI)-->.

@@

To gain insight to the eco-evolutionary feedback generated by character displacement, I conducted simulations using an Adaptive Dynamics approach. 
Specifically, after letting consumer and resource abundances reach a steady state, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability. 
The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006]. 
This function has the useful property that it differentiates between cases where intermediate combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ are higher than the extremes (when $n>1$, green line in Fig. \ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line in Fig. \ref{fig:tradeoff}). 
When $n=1$, the tradeoff function is linear, and all combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ have the same total attack rate (blue line in Fig. \ref{fig:tradeoff}). 
Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource abundances to reach a steady state. 
I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0). 
Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and $w_{11} = w_{22} = 0.6$. 
I set an initial value of $a_{11} = a_{22} = 1.2$, while $a_{12}$ and $a_{21}$ depended on the value of *n*. 
I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model)<!-- and has been archived with Zenodo (INSERT DOI)-->.

>>>>

*Data accessibility*: No new data were used. All code used to reproduce this manuscript and simulations within are publically available on GitHub (https://github.com/mabarbour/ECD_model) and has been archived with Zenodo (https://zenodo.org/badge/latestdoi/79349192).

@@

To gain insight to the eco-evolutionary feedback generated by character displacement, I conducted simulations using an Adaptive Dynamics approach. 
Specifically, after letting consumer and resource abundances reach a steady state, I created a mutant consumer by randomly choosing one and modifying its attack rate on one resource by either subtracting or adding a small constant (0.01 in the following simulations) with equal probability. 
The mutant's attack rate on the other resource was determined by a tradeoff, such that $\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1$, where $A$ is the total investment in attack rates and $n$ describes the shape of the tradeoff [@Sargent2006]. 
This function has the useful property that it differentiates between cases where intermediate combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ are higher than the extremes (when $n>1$, green line in Fig. \ref{fig:tradeoff}) or, conversely, where the two extremes are higher than intermediate investments (when $n<1$, orange line in Fig. \ref{fig:tradeoff}). 
When $n=1$, the tradeoff function is linear, and all combinations of $a_{\text{1},j}$ and $a_{\text{2},j}$ have the same total attack rate (blue line in Fig. \ref{fig:tradeoff}). 
Assuming the mutant consumer was rare, I then determined whether the mutant had higher relative fitness than the resident consumer, and thus could invade and replace the resident consumer. 
If the mutant was able to invade, I updated the attack rate of the resident consumer to the mutant attack rate and allowed consumer and resource abundances to reach a steady state. 
I then repeated the simulation up to 10,000 times, which was sufficient for consumers to either reach an evolutionary stable strategy [ESS, @Smith1973] or an evolutionary limit (e.g., $\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}$ is constrained to a maximum of 1 and minimum of 0). 
Unless otherwise noted, I conducted simulations with the following parameter values: *r* = 1; *K* = 4; *e* = 0.8; *m* = 1; *A* = 2; *h* = 0.4; and $w_{11} = w_{22} = 0.6$. 
I set an initial value of $a_{11} = a_{22} = 1.2$, while $a_{12}$ and $a_{21}$ depended on the value of *n*. 
I set initial consumer and resource abundances to: $R_1 = R_2 = 2$; $C_1 = C_2 = 1$.
All simulations were conducted in R [@R] and the code to reproduce these simulations is publically available on GitHub (https://github.com/mabarbour/ECD_model) and has been archived with Zenodo (https://zenodo.org/badge/latestdoi/79349192).

### 2019-11-18 manuscript_ECD_model.tex
added Zenodo DOI to manuscript.

\emph{Data accessibility}: No new data were used. All code used to
reproduce this manuscript and simulations within are publically
available on GitHub (\url{https://github.com/mabarbour/ECD_model}).

@@

To gain insight to the eco-evolutionary feedback generated by character
displacement, I conducted simulations using an Adaptive Dynamics
approach. Specifically, after letting consumer and resource abundances
reach a steady state, I created a mutant consumer by randomly choosing
one and modifying its attack rate on one resource by either subtracting
or adding a small constant (0.01 in the following simulations) with
equal probability. The mutant's attack rate on the other resource was
determined by a tradeoff, such that
\(\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1\),
where \(A\) is the total investment in attack rates and \(n\) describes
the shape of the tradeoff \citep{Sargent2006}. This function has the
useful property that it differentiates between cases where intermediate
combinations of \(a_{\text{1},j}\) and \(a_{\text{2},j}\) are higher
than the extremes (when \(n>1\), green line in Fig. \ref{fig:tradeoff})
or, conversely, where the two extremes are higher than intermediate
investments (when \(n<1\), orange line in Fig. \ref{fig:tradeoff}). When
\(n=1\), the tradeoff function is linear, and all combinations of
\(a_{\text{1},j}\) and \(a_{\text{2},j}\) have the same total attack
rate (blue line in Fig. \ref{fig:tradeoff}). Assuming the mutant
consumer was rare, I then determined whether the mutant had higher
relative fitness than the resident consumer, and thus could invade and
replace the resident consumer. If the mutant was able to invade, I
updated the attack rate of the resident consumer to the mutant attack
rate and allowed consumer and resource abundances to reach a steady
state. I then repeated the simulation up to 10,000 times, which was
sufficient for consumers to either reach an evolutionary stable strategy
\citep[ESS,][]{Smith1973} or an evolutionary limit (e.g.,
\(\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}\) is constrained to a
maximum of 1 and minimum of 0). Unless otherwise noted, I conducted
simulations with the following parameter values: \emph{r} = 1; \emph{K}
= 4; \emph{e} = 0.8; \emph{m} = 1; \emph{A} = 2; \emph{h} = 0.4; and
\(w_{11} = w_{22} = 0.6\). I set an initial value of
\(a_{11} = a_{22} = 1.2\), while \(a_{12}\) and \(a_{21}\) depended on
the value of \emph{n}. I set initial consumer and resource abundances
to: \(R_1 = R_2 = 2\); \(C_1 = C_2 = 1\). All simulations were conducted
in R \citep{R} and the code to reproduce these simulations is publically
available on GitHub (\url{https://github.com/mabarbour/ECD_model}).

>>>>

\emph{Data accessibility}: No new data were used. All code used to
reproduce this manuscript and simulations within are publically
available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
has been archived with Zenodo
(\url{https://zenodo.org/badge/latestdoi/79349192}).

@@

To gain insight to the eco-evolutionary feedback generated by character
displacement, I conducted simulations using an Adaptive Dynamics
approach. Specifically, after letting consumer and resource abundances
reach a steady state, I created a mutant consumer by randomly choosing
one and modifying its attack rate on one resource by either subtracting
or adding a small constant (0.01 in the following simulations) with
equal probability. The mutant's attack rate on the other resource was
determined by a tradeoff, such that
\(\big(a_{\text{1},j}/{A}\big)^n+\big(a_{\text{2},j}/{A}\big)^n=1\),
where \(A\) is the total investment in attack rates and \(n\) describes
the shape of the tradeoff \citep{Sargent2006}. This function has the
useful property that it differentiates between cases where intermediate
combinations of \(a_{\text{1},j}\) and \(a_{\text{2},j}\) are higher
than the extremes (when \(n>1\), green line in Fig. \ref{fig:tradeoff})
or, conversely, where the two extremes are higher than intermediate
investments (when \(n<1\), orange line in Fig. \ref{fig:tradeoff}). When
\(n=1\), the tradeoff function is linear, and all combinations of
\(a_{\text{1},j}\) and \(a_{\text{2},j}\) have the same total attack
rate (blue line in Fig. \ref{fig:tradeoff}). Assuming the mutant
consumer was rare, I then determined whether the mutant had higher
relative fitness than the resident consumer, and thus could invade and
replace the resident consumer. If the mutant was able to invade, I
updated the attack rate of the resident consumer to the mutant attack
rate and allowed consumer and resource abundances to reach a steady
state. I then repeated the simulation up to 10,000 times, which was
sufficient for consumers to either reach an evolutionary stable strategy
\citep[ESS,][]{Smith1973} or an evolutionary limit (e.g.,
\(\frac{a_{ij}}{a_{\text{1},j}+a_{\text{2},j}}\) is constrained to a
maximum of 1 and minimum of 0). Unless otherwise noted, I conducted
simulations with the following parameter values: \emph{r} = 1; \emph{K}
= 4; \emph{e} = 0.8; \emph{m} = 1; \emph{A} = 2; \emph{h} = 0.4; and
\(w_{11} = w_{22} = 0.6\). I set an initial value of
\(a_{11} = a_{22} = 1.2\), while \(a_{12}\) and \(a_{21}\) depended on
the value of \emph{n}. I set initial consumer and resource abundances
to: \(R_1 = R_2 = 2\); \(C_1 = C_2 = 1\). All simulations were conducted
in R \citep{R} and the code to reproduce these simulations is publically
available on GitHub (\url{https://github.com/mabarbour/ECD_model}) and
has been archived with Zenodo
(\url{https://zenodo.org/badge/latestdoi/79349192}).
